QALY Comparison - Living Wage Intervention

### 2023-11-01 minos/testing/QALY_comparison_energyCrises.Rmd
Updated the scripts for QALY comparison of living wage and energy crisis interventions

require(tidyverse)
require(ggplot2)
require(knitr)
require(here)
require(MESS)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
)
rm(workingDir)
```

@@

ggplot(data = combined, aes(x = year, y = QALYs, group = intervention, colour = intervention)) +
  geom_line()
```

@@

```{r}
#base.auc <- sum(diff(base$time)*rollmean(base$QALYs, 2))
base.auc <- data.frame(intervention = 'baseline', AUC = auc(x=base$year, y=base$QALYs, type = 'spline'))
support.auc <- data.frame(intervention = 'Support', AUC = auc(x=support$year, y=support$QALYs, type = 'spline'))
noSupport.auc <- data.frame(intervention = 'NoSupport', AUC = auc(x=noSupport$year, y=noSupport$QALYs, type = 'spline'))

@@

ggplot(data = combined.auc, aes(x = intervention, y = AUC, group = intervention, fill = intervention)) +
  geom_col()
```

```{r}
# Another version showing difference from baseline for better comparison
combined.auc2 <- combined.auc %>%
  pivot_wider(names_from = 'intervention',
              values_from = 'AUC') %>%
  mutate(change_baseline = baseline - baseline,
         change_Support = Support - baseline,
         change_noSupport = NoSupport - baseline) %>%
  pivot_longer(cols = change_baseline:change_noSupport,
               names_to = 'intervention',
               names_prefix = 'change_',
               values_to = 'AUC_difference')

@@

ggplot(combined.auc2, aes(x = intervention, y = AUC_difference, group = intervention, fill = intervention)) +
  geom_col()
```

>>>>

require(tidyverse)
require(ggplot2)
require(knitr)
require(here)
require(MESS)
require(scales)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
  options(scipen = 999)
)
rm(workingDir)
```

@@

ggplot(data = combined, aes(x = year, y = QALYs, group = intervention, colour = intervention)) +
  geom_smooth()
```

@@

```{r}
base.auc <- base %>%
  group_by(run_id) %>%
  summarise(AUC = auc(x=year, y = QALYs, type = 'spline')) %>%
  mutate(intervention = 'baseline')

support.auc <- support %>%
  group_by(run_id) %>%
  summarise(AUC = auc(x=year, y = QALYs, type = 'spline')) %>%
  mutate(intervention = 'Support')

noSupport.auc <- noSupport %>%
  group_by(run_id) %>%
  summarise(AUC = auc(x=year, y = QALYs, type = 'spline')) %>%
  mutate(intervention = 'NoSupport')

@@

ggplot(data = combined.auc, aes(x = intervention, y = AUC, group = intervention, fill = intervention)) +
  stat_summary(fun = 'mean', geom = 'bar') +
  geom_errorbar(aes(ymin = min(AUC), ymax = max(AUC), width = 0.4)) +
  labs(title = 'Total AUC comparison', subtitle = 'Baseline vs Energy Crisis Interventions') +
  xlab('Intervention')
```

```{r}
# Another version showing difference from baseline for better comparison
combined.auc2 <- combined.auc %>%
  pivot_wider(names_from = 'intervention',
              values_from = c('AUC')) %>%
  mutate(change_baseline = baseline - baseline,
         change_Support = Support - baseline,
         change_noSupport = NoSupport - baseline) %>%
  pivot_longer(cols = change_baseline:change_noSupport,
               names_to = 'intervention',
               names_prefix = 'change_',
               values_to = 'AUC_difference') %>%
  filter(intervention != 'baseline')

@@

ggplot(combined.auc2, aes(x = intervention, y = AUC_difference, group = intervention, fill = intervention)) +
  stat_summary(fun = 'mean', geom = 'bar') +
  geom_errorbar(aes(ymin = min(AUC_difference), ymax = max(AUC_difference), width = 0.4)) +
  labs(title = 'Change in AUC', subtitle = 'Baseline vs Energy Crisis Interventions') +
  xlab('Intervention')
```

# Cost per QALY

The [UK Government](https://www.gov.uk/government/publications/valuation-of-risks-to-life-and-health-monetary-value-of-a-life-year-voly/a-scoping-study-on-the-valuation-of-risks-to-life-and-health-the-monetary-value-of-a-life-year-voly) has set the value of 1 QALY to be equal to $60,000. This is the number we will use for our calculations.

```{r}
combined.cost <- combined %>%
  select(-pop_size, -SF_12_MCS, -SF_12_PCS, -utility) %>%
  mutate(QALY_value = QALYs * 60000)

ggplot(combined.cost, aes(x = year, y = QALY_value, group = intervention, colour = intervention)) +
  geom_smooth() +
  scale_y_continuous(label = comma)

combined.cost.change <- combined.cost %>%
  select(-total_boost, -QALYs) %>%
  pivot_wider(names_from = 'intervention',
              values_from = 'QALY_value') %>%
  mutate(change_baseline = baseline - baseline,
         change_Support = energyDownlift - baseline,
         change_NoSupport = energyDownliftNoSupport - baseline) %>%
  select(-baseline, -energyDownlift, -energyDownliftNoSupport) %>%
  pivot_longer(cols = change_baseline:change_NoSupport,
               names_to = 'intervention',
               names_prefix = 'change_',
               values_to = 'QALY_value_change') %>%
  filter(intervention != 'baseline')

ggplot(combined.cost.change, aes(x = year, y = QALY_value_change, group = intervention, colour = intervention, fill = intervention)) +
  geom_smooth() +
  scale_y_continuous(label = comma)
```

```{r}
total.combined.cost <- combined.cost.change %>%
  group_by(run_id, intervention) %>%
  summarise(TotalChange = sum(QALY_value_change))

ggplot(total.combined.cost, aes(x = intervention, y = TotalChange, group = intervention, fill = intervention)) +
  stat_summary(fun = 'mean', geom = 'bar') +
  geom_errorbar(aes(ymin = min(TotalChange), ymax = max(TotalChange), width = 0.4)) +
  labs(title = 'Total Change in QALY Value', subtitle = 'Baseline vs Energy Crisis Interventions') +
  xlab('Intervention')
```

# Incremental Cost-Effectiveness Ratio

From [NICE guidelines](https://www.nice.org.uk/process/pmg6/chapter/assessing-cost-effectiveness) about the cost-effectiveness of health interventions (7.3), and works with a statistic called the Incremental Cost-Effectiveness Ratio (ICER). The formula for calculating the ICER can be found [here](https://en.wikipedia.org/wiki/Incremental_cost-effectiveness_ratio), but in short, it represents the average incremental cost associated with 1 additional unit of the measurement of effect. In our case, the unit of measurement of effect is a QALY.

From the NICE guidelines, there is a general consensus that anything below £20,000 is cost effective, and anything above £30,000 is not. Between £20,000 and £30,000 is a grey area, and depends on the field and other factors to decide (i.e. if a rare or important condition is tackled by an intervention then the cost-effective threshold is higher).

```{r}
combined.cost.mean <- combined.cost %>%
  group_by(run_id, year, intervention) %>%
  summarise(across(everything(), mean))

# ICER = (C1 - C0) / (E1 - E0)
# where
# C1 = cost of intervention
# E1 = effect of intervention
# C0 = cost of control group
# E0 = effect of control group

ICER <- combined.cost.mean %>%
  pivot_wider(names_from = 'intervention',
              values_from = c('total_boost', 'QALYs', 'QALY_value')) %>%
  mutate(ICER_Support = (total_boost_energyDownlift - total_boost_baseline) / (QALYs_energyDownlift - QALYs_baseline),
         ICER_NoSupport = (total_boost_energyDownliftNoSupport - total_boost_baseline) / (QALYs_energyDownliftNoSupport - QALYs_baseline)) %>%
  select(run_id, year, ICER_Support, ICER_NoSupport) %>%
  pivot_longer(cols = -c(run_id, year),
               names_prefix = 'ICER_',
               names_to = 'intervention',
               values_to = 'ICER')

ggplot(ICER, aes(x = year, y = ICER, group = intervention, fill = intervention, colour = intervention)) +
  geom_smooth()

ICER.final <- ICER %>%
  group_by(run_id, intervention) %>%
  summarise(ICER = mean(ICER))

ggplot(ICER.final, aes(x = intervention, y = ICER, group = intervention, fill = intervention)) +
  stat_summary(fun = 'mean', geom = 'bar') +
  geom_errorbar(aes(ymin = min(ICER), ymax = max(ICER), width = 0.4)) +
  labs(title = 'ICER', subtitle = 'Living Wage Intervention') +
  xlab('Intervention')
```

```{r}

```

### 2023-11-01 minos/testing/QALY_comparison_livingWage.Rmd
Updated the scripts for QALY comparison of living wage and energy crisis interventions

combined.auc <- rbind(base.auc, wage.auc)

# #base.auc <- sum(diff(base$time)*rollmean(base$QALYs, 2))
# base.auc <- data.frame(intervention = 'baseline', AUC = auc(x=base$year, y=base$QALYs, type = 'spline'))
# support.auc <- data.frame(intervention = 'Support', AUC = auc(x=support$year, y=support$QALYs, type = 'spline'))
# noSupport.auc <- data.frame(intervention = 'NoSupport', AUC = auc(x=noSupport$year, y=noSupport$QALYs, type = 'spline'))
# 
# combined.auc <- rbind(base.auc, support.auc, noSupport.auc)
```

@@

```{r}

# pivot_longer(cols = change_baseline:change_LivingWage,
#                names_to = 'intervention',
#                names_prefix = 'change_',
#                values_to = 'AUC_difference') %>%
#   filter(intervention != 'baseline')

combined.cost <- combined %>%
  select(-pop_size, -SF_12_MCS, -SF_12_PCS, -utility) %>%
  mutate(QALY_value = QALYs * 60000)

@@

ggplot(combined.cost.change, aes(x = year, y = QALY_value_change, group = intervention, fill = intervention)) +
  geom_smooth() +
  scale_y_continuous(label = comma)
```

>>>>

combined.auc <- rbind(base.auc, wage.auc)
```

@@

```{r}
combined.cost <- combined %>%
  select(-pop_size, -SF_12_MCS, -SF_12_PCS, -utility) %>%
  mutate(QALY_value = QALYs * 60000)

@@

ggplot(combined.cost.change, aes(x = year, y = QALY_value_change, group = intervention, colour = intervention, fill = intervention)) +
  geom_smooth() +
  scale_y_continuous(label = comma)
```
